---
title: "PPOL561 - HM 1 - Data Wrangling and Visualization"
author: "Allen Church"
output: pdf_document
---
```{r,include=F}
knitr::opts_chunk$set(warning = F,error=F,message=F)
require(tidyverse)
```
## Question 2
1. Only countries are in the data; no regions or non-state entities
2. Drop all variables but the `iso3c`,`date`,`value`, `indicatorID`, and `country`.
3. Spread `indicatorID` into two new variables ("SP.POP.TOTL","NY.GDP.MKTP.CD") where the values correspond with the entries in the `value` variable column. 
4. Rename `iso3c` to `ccode`, `date` to `year`, `SP.POP.TOTL` to `pop`, and `NY.GDP.MKTP.CD` to `gdp`.
5. Rearrange the columns in the following order: `country`, `ccode`, `year`, `gdp`, `pop`
6. Count to see if you have any missing for `gdp` entries; if there are any missing entries, create a new variable called `missing_gdp` that takes on the value of `1` if missing, `0` otherwise. How many country-year values are missing? Which country has the most missing values?
  - tip: look at the `ifelse()` and `is.na()` functions for ideas regarding how to do this.
7. Transform `pop` and `gdp` using the natural `log()`. Name the new variables `ln_gdp` and `ln_pop`.
8. Export the data as a `.dta` and `.csv` into the `Output/` folder. Name the data `wb-pop-gdp-2000-2005`.

Please use the tidyverse function to perform _all_ these tasks. Be sure to pipe (` %>% `) when necessary.

### Part 1: Import population and GDP data as dataframe, specifying regions and dates:
```{r}
library(wbstats)
library(ggthemes)

#Call regions, specify PPG indicator, and start/end dates
# indicator = External debt stocks, public and publicly guaranteed
ext_debt <- wb(country = c("EAP", "MNA", "ECA", "SAS", "LAC", "SSA"), indicator = "DT.DOD.DPPG.CD",
              startdate = 2008, enddate = 2018)

head(ext_debt)
```

### Parts 2-5: Initial data wrangling - Adjusting columns
```{r}
#Select columns from dataframe above, spread data from horizontal to vertical format, rename columns, mutate to convert units to billions, and select final columns
ext_debt <- ext_debt %>%  
  select("iso3c","date","value", "indicatorID") %>% 
  spread(key = indicatorID, value = value) %>% 
  rename(region = iso3c,
         year = date,
         ppg = DT.DOD.DPPG.CD) %>%
  mutate(PPG_in_billions = ppg/1000000000) %>%
  select("region", "year", "PPG_in_billions")

head(ext_debt)
```

```{r}
#remove grid and rename regions
ggplot(ext_debt, aes(x=year, y=PPG_in_billions, color = region, group = region)) +
  geom_point() + geom_line() +
  labs(title="External debt stocks: Public and Publicly Guaranteed (PPG)",
  subtitle = "Long-term external obligations of public debtors",
  caption = "Source: International Debt Statistics",
  x="Year", y="USD Billions") +
  theme_minimal()
```

Call External Debt to GNI: 2008 - 2018 for all regions
```{r}
#Call API, select external debt stocks as % of GNI, and rename columns
debt_gni <- wb(country = c("EAP", "MNA", "ECA", "SAS", "LAC", "SSA"), indicator = "DT.DOD.DECT.GN.ZS",
              startdate = 2008, enddate = 2018) %>% 
  select("iso3c","date","value", "indicatorID") %>% 
  spread(key = indicatorID, value = value) %>% 
  rename(region = iso3c,
         year = date,
         debt_to_gni = DT.DOD.DECT.GN.ZS) %>%
  select("region", "year", "debt_to_gni")
  
head(debt_gni)
```

Plotting External Debt Stocks to GNI (2008 - 2018)
```{r}
test_graph <- ggplot(debt_gni, aes(x=year, y=debt_to_gni, color = region, group = region)) +
  geom_point() + geom_line() +
  labs(title="Total external debt stocks to Gross National Income (GNI)",
  caption = "Source: International Debt Statistics",
  x="Year", y="Percent of GNI") +
  theme_minimal()

test_plot <- ggplotly(test_graph)

test_plot
```


```{r}
library(maps)

map_data("world")

#Filter by only capital cities
world.cities %>% filter(capital==1)

head(world.cities)
```

Relationship between Exports of Goods (BX.GSR.TOTL.CD) and Commitments (DT.COM.DPPG.CD)
```{r}
export_com <- wb(country = "LAC", indicator = c("BX.GSR.TOTL.CD", "DT.COM.DPPG.CD"),
              startdate = 2008, enddate = 2018) %>% 
  select("iso3c","date","value", "indicatorID", "country") %>% 
  spread(key = indicatorID, value = value) %>% 
  rename(region = iso3c,
         year = date,
         export = BX.GSR.TOTL.CD,
         com = DT.COM.DPPG.CD) %>% 
  mutate(Export_in_billions = export/1000000000) %>% 
  mutate(Commitments_in_billions = com/1000000000) %>%
  select("country", "region", "year", "Export_in_billions", "Commitments_in_billions")
```

Plotting
```{r}
ggplot() +
  geom_line(data = export_com, aes(x = year, y = Export_in_billions, group = region), color="blue") +
  geom_point() + 
  geom_line(data = export_com, aes(x = year, y = Commitments_in_billions, group = region), color = "red") +
  labs(title="Total Commitments to Exports",
  caption = "Source: International Debt Statistics",
  x="Year", y="Billions") +
  theme_minimal()
```


```{r}
data %>% 
  #filter(continent=="Asia") %>% 
  mutate(ln_pop = log(pop),
         ln_gdp = log(gdpPercap)) %>% 
  ggplot(aes(x = ln_gdp, y = lifeExp, size=ln_pop,color=continent)) +
  geom_point(alpha=.5) +
  labs(y="Life Expectancy",x="Log GDP", size="Log Population") +
  #se=F sets standard error to false
  geom_smooth(se=F) +
  theme(legend.position = "bottom") +
  ggthemes::scale_color_gdocs() +
  #facet wrap and ~ creates separate graphs for each ~continent in this case
  facet_wrap(~continent, nrow = 1)
```

